Variations on a Bayesian Theme: Comparing Bayesian Models of Referential Reasoning

### 2014-03-15 02_RSA.tex
intro to FG's experiment

Now let us turn to play as the listener. If we hear ``Square'' or
``Blue'', we easily know the intended referent as there is no
ambiguity, but what if we hear ``Circle'' (or ``Green'')? There are
two circles in the context that we need to choose from. On the one
hand, the blue circle, having the unique color in the context, seems
to be perceptually dominant and thus easily captures our attention. On
the other hand, from the previous analysis we know that if the blue
circle were the intended referent, the speaker could have chosen
``Blue'' which is not ambiguous and thus more informative. Hence the
listener needs to balance two sources of information, i.e., the
(presumably subconscious) perceptual salience of different objects and
the rational expectation of the likelihood of the speaker making the
utterance for each object. The latter line of reasoning is crucial to
the classic Gricean account of scalar implicatures, and the major
challenges in terms of formal modeling are how to quantify notions
such as informativeness and how different pieces of information should
be integrated.\todo{discuss Frank and Goodman's experiments and
  results (in as far possible) here, before venturing into the model}

>>>>

Now let us turn to play as the listener. If we hear ``Square'' or
``Blue'', we easily know the intended referent as there is no
ambiguity, but what if we hear ``Circle'' (or ``Green'')? There are
two circles in the context that we need to choose from. On the one
hand, the blue circle, having the unique color in the context, seems
to be perceptually dominant and thus easily captures our attention. On
the other hand, from the previous analysis we know that if the blue
circle were the intended referent, the speaker could have chosen
``Blue'' which is not ambiguous and thus more informative. Hence the
listener needs to balance two sources of information, i.e., the
(presumably subconscious) perceptual salience of different objects and
the rational expectation of the likelihood of the speaker making the
utterance for each object.

The experiment in \cite{Frank} confirmed the above intuitions about the 
referential games. 
In their study, 745 US participants were recruited via 
Amazon’s Mechanical Turk to take part in a one-shot referential game, either 
as the speaker or the listener. 
The context of the game always consisted of three objects. 
Each object had three features: color (red, blue, or green), shape 
(circle, square, or cloud), and texture (solid, or polka-dot, or striped).
Every context had one feature fixed for all objects and the other two features
systematically varied. 
The objects were randomly permuted and the target object was randomly assigned. 
Depending on how many objects share the same value of a feature with the target, 
the contexts can be classified into 7 categories: 1/1, 1/2 (the target's 
first feature is unique, and there is another object sharing the same value of
the second feature with the target), 1/3, 2/2 with the same object sharing both
features with the target (which means there is another object identical to the 
target), 2/2 with different objects sharing the features (such as in 
Fig.~\ref{context}), 2/3 and 3/3. 
The speakers were asked to choose a feature to refer to the target object. 
In order to measure the perceptual salience, some of the listeners were told 
that the speaker was using a language that they did not understand and then 
they were asked to guess the intended referent. 
Other listeners were told a feature of the target and asked to choose the 
intended referent.

The result of the experiment showed that speakers generally followed 
Grice's Maxim of Quantity, i.e., they tried to be informative by choosing the 
most unique feature of the target, and the listeners were influenced by both 
the perceptual salience and the informativeness of the utterance that the 
speaker produced. 
However, the data, being quantitative in nature, are also full of apparent 
exceptions. 
This means that the empirical data cannot be fully explained by the classic 
Gricean account of scalar implicatures, and the major challenges in terms of 
formal modeling include how to quantify notions
such as informativeness and how different pieces of information should
be integrated.
